---
title: "P8105 Homework 1"
author: "ss7636"
date: "9/19/25"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r}
library(moderndive)
library(tidyverse)
data("early_january_weather")
early_january_weather
```
```{r}
n_row = nrow(early_january_weather) #calculates number of rows
n_col = ncol(early_january_weather) #calculates number of columns
mean_temp = mean(pull(early_january_weather, temp)) #calculates the mean of the temperature column
```

The dataset has `r n_row` rows and `r n_col` columns.
he dataset contains the following variables: `r paste(names(early_january_weather), collapse = ", ")`.
The important variables are:
`temp`: temperature in Ferenheit
`month`: 1 (January)
`year`: 2013
`humid`: humidity
The mean temperature is `r mean_temp` degrees Ferenheit. 

```{r}
library(ggplot2)
plot = ggplot(early_january_weather, aes(x= time_hour, y = temp, color= humid)) +
  geom_point()+
  labs(title = "Temperature vs Hour Colored by Humidity", 
       x = "Hour in January",
       y = "Temperature in Farenheit")

plot
```
```{r}
ggsave("weather_plot.png", plot = plot)
```

From January 1st through Janurary 14th,, Temperatures rise in the day and drop at night. As January progresses (after January 7th), temperature rises overall through the day for both morning, afternoons, and nights compared to before January 7th. Humidity is higher after January 7th than before January 7th. 

## Problem 2


```{r}
df = 
  tibble(
num = rnorm(10),            
logical = num > 0,          
char=  letters[1:10],            
factor = factor(sample(c("Small", "Medium", "Large"), 10, replace = TRUE))
)

#The following lines of codes calculates the mean of the variables we created. 
mean(pull(df, num))        
mean(pull(df, logical))    
mean(pull(df, char))          
mean(pull(df, factor))  

```

Cannot take the mean of a character or factor. We were able to take the mean of numeric and logical variables.  

```{r}
#The following lines of codes converts the variable to numeric.
as.numeric(pull(df, logical)) 
as.numeric(pull(df, char)) 
as.numeric(pull(df, factor)) 
```
For logical vectors, TRUE values are converted to 1 and FALSE values are converted to 0. This makes sense because R internally stores logicals as binary values. Taking the mean of a logical vector works because it gives the proportion of TRUEs. This does not work for character vectors and we get NA values with a warning because arbitrary characters cannot be meaningfully mapped to numbers. As a result, the mean cannot be calculated.Lastly, for the factor, we see the values that the factor interanally stores the levels as. For example, the first level is coded as 1, the second as 2, and so on. as.numeric() reveals those codes. While we can technically compute a mean of these numbers, it is not meaningful because the numeric codes don’t represent real quantitative values. This helps explain why the mean function only gives us a meaningful output for the numerical and logical values, and why it does not give us anything for the character and factor values. 


